\documentclass[a4paper]{article}

\title{First-Order Logic According to Harrison}

\author{Alexander Birch Jensen, Anders Schlichtkrull \&\\ J{\o}rgen Villadsen, DTU Compute, Denmark}

\date{\isadate\today}

\usepackage[left=15mm,right=15mm,top=20mm,bottom=27mm]{geometry}

\usepackage{datetime,isabelle,isabellesym,parskip}

\newdateformat{isadate}{\THEDAY\ \monthname[\THEMONTH] \THEYEAR}

\usepackage{pdfsetup}

\isabellestyle{tt}

\urlstyle{rm}

%\renewcommand{\isachardoublequote}{}
%\renewcommand{\isachardoublequoteopen}{}
%\renewcommand{\isachardoublequoteclose}{}

\renewcommand{\isamarkupchapter}[1]{\clearpage\isamarkupsection{#1}\medskip}
\renewcommand{\isamarkupsection}[1]{\medskip\section*{#1}\addcontentsline{toc}{section}{#1}\medskip}
\renewcommand{\isamarkupsubsection}[1]{\medskip\subsection*{#1}\medskip}
\renewcommand{\isamarkupsubsubsection}[1]{\medskip\subsubsection*{#1}\medskip}

\renewcommand{\isabeginpar}{\par\ifisamarkup\relax\else\bigskip\fi}
\renewcommand{\isaendpar}{\par\bigskip}

\begin{document}

\makeatletter
\parbox[t]{\textwidth}{\centering\Huge\bfseries\@title}\par\kern5mm
\parbox[t]{\textwidth}{\centering\Large\bfseries\@author}\par\kern3mm
\parbox[t]{\textwidth}{\centering\bfseries\@date}\par\kern8mm
\makeatother

\begin{abstract}\normalsize\noindent
We present a certified declarative first-order prover with equality based on John Harrison's
Handbook of Practical Logic and Automated Reasoning, Cambridge University Press, 2009. ML code
reflection is used such that the entire prover can be executed within Isabelle as a very simple
interactive proof assistant. As examples we consider Pelletier's problems 1-46.
\end{abstract}

\tableofcontents

\isamarkupsection{Preample}

Preliminary formalizations are described here:
\begin{trivlist}
\item
Alexander Birch Jensen:
\emph{Development and Verification of a Proof Assistant.}
Master's Thesis, Technical University of Denmark, 2016.
\url{http://findit.dtu.dk/en/catalog/2345011633}
\item
Alexander Birch Jensen, Anders Schlichtkrull \&\ J{\o}rgen Villadsen:
\emph{Verification of an LCF-Style First-Order Prover with Equality.}
Isabelle Workshop 2016.
\url{https://github.com/logic-tools/sml-handbook}
\end{trivlist}

\clearpage

\input{session}

\clearpage

\isamarkupsection{Acknowledgements}

The SML code is based on the OCaml code accompanying John Harrison's
Handbook of Practical Logic and Automated Reasoning, Cambridge University Press, 2009

Thanks to Jasmin Blanchette, Andreas Halkj{\ae}r From, John Bruntse Larsen, Andrei Popescu and
Tom Ridge for discussions.

\end{document}
