\documentclass[11pt,a4paper,notitlepage]{report}
\usepackage{isabelle,isabellesym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{wasysym}
\usepackage{graphicx}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\renewcommand{\setisabellecontext}[1]{\markright{#1}}

\begin{document}

\title{Towards Certified Slicing}
\author{Daniel Wasserrab}
\maketitle

\begin{abstract}
Slicing is a widely-used technique with applications in e.g. compiler
technology and software security. Thus verification of
algorithms in these areas is often based on the correctness of slicing,
which should ideally be proven independent of concrete programming
languages and with the help of well-known verifying techniques such as
proof assistants. As a first step in this direction, this contribution
presents a framework for dynamic \cite{WasserrabL:08} 
and static intraprocedural slicing \cite{WasserrabLS:09}
based on control flow and program dependence graphs.
Abstracting from concrete syntax we base the framework on a graph
representation of the program fulfilling certain structural
and well-formedness properties.

We provide two instantiations to show the validity of the framework:
a simple While language and the sophisticated object-oriented byte code
language from Jinja \cite{KleinN:06}. 
\end{abstract}

\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

\begin{thebibliography}{10}

\bibitem{WasserrabLS:09}
Daniel Wasserrab and Denis Lohner and Gregor Snelting
\newblock On PDG-Based Noninterference and its Modular Proof
\newblock In {\em Proc. of PLAS'09}, pages 31--44. ACM, 2009.

\bibitem{WasserrabL:08}
Daniel Wasserrab and Andreas Lochbihler.
\newblock Formalizing a framework for dynamic slicing of program dependence
  graphs in {Isabelle/HOL}.
\newblock In {\em Proc. of TPHOLS'08}, pages 294--309. Springer-Verlag, 2008.

\bibitem{KleinN:06}
Gerwin Klein and Tobias Nipkow.
\newblock {A Machine-Checked Model for a Java-Like Language, Virtual Machine
  and Compiler}.
\newblock {\em ACM Transactions on Programming Languages and Systems},
  28(4):619--695, 2006.

\end{thebibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
