\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

\usepackage{amssymb}
\usepackage{xspace}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\newcommand\isafor{\textsf{IsaFoR}}
\newcommand\ceta{\textsf{Ce\kern-.18emT\kern-.18emA}}
\newcommand\rats{\mathbb{Q}}
\newcommand\ints{\mathbb{Z}}
\newcommand\nats{\mathbb{N}}
\newcommand\reals{\mathbb{R}}
\newcommand\mod{\mathit{mod}}
\newcommand\complex{\mathbb{C}}

\newcommand\rai{real algebraic number\xspace}
\newcommand\rais{real algebraic numbers\xspace}

\begin{document}

\title{Polynomial Factorization\footnote{Supported by FWF (Austrian Science Fund) project Y757.}}
\author{Ren\'e Thiemann and Akihisa Yamada}
\maketitle

\begin{abstract}
Based on existing libraries for polynomial interpolation and matrices,
we formalized several factorization algorithms for polynomials, including
Kronecker's algorithm for integer polynomials,
Yun's square-free factorization algorithm for field polynomials, and
a factorization algorithm which delivers root-free polynomials. 

As side products, we developed division algorithms for polynomials over integral domains,
as well as primality-testing and prime-factorization algorithms for integers.
\end{abstract}

\tableofcontents

\section{Introduction}

The details of the factorization algorithms have mostly been extracted 
from Knuth's Art of Computer Programming
\cite{Knuth}. Also Wikipedia provided valuable help.

\medskip
As a first fast
preprocessing for factorization we integrated Yun's factorization algorithm which identifies duplicate
factors \cite{Yun}. In contrast to the existing formalized result that the GCD of $p$ and $p'$ has no
duplicate factors (and the same roots as $p$), Yun's algorithm decomposes a polynomial $p$ into
$p_1^1 \cdot \ldots \cdot p_n^n$ such that no $p_i$ has a duplicate factor and there is no common
factor of $p_i$ and $p_j$ for $i \neq j$. As a comparison, the GCD of $p$ and $p'$ is exactly
$p_1 \cdot \ldots \cdot p_n$, but without decomposing this product into the list of $p_i$'s.

Factorization over $\rats$ is reduced to factorization over $\ints$ 
with the help of Gauss' Lemma.

Kronecker's algorithm for factorization over $\ints$ requires both
polynomial interpolation over $\ints$ and prime factorization over $\nats$. Whereas the former
is available as a separate AFP-entry, for prime factorization we mechanized a simple algorithm depicted
in \cite{Knuth}:
For a given number $n$,
the algorithm iteratively checks divisibility by numbers until $\sqrt n$,
with some optimizations:
it uses a precomputed set of small primes (all primes up to 1000), 
and if $n\ \mod\ 30 = 11$, the next test candidates in the range $[n,n+30)$ 
are only the 8 numbers $n,n+2,n+6,n+8,n+12,n+18,n+20,n+26$.

However, in theory and praxis it turned out that Kronecker's algorithm is too inefficient. 
Therefore, in a separate AFP-entry we formalized the Berlekamp-Zassenhaus 
factorization.\footnote{The Berlekamp-Zassenhaus AFP-entry was originally not present 
and at that time,
this AFP-entry contained an implementation of Berlekamp-Zassenhaus as a
non-certified function.}


There also is a combined factorization algorithm: For polynomials of degree 2, the 
closed form for the roots of quadratic polynomials is applied. For polynomials of degree 3, 
the rational root test determines whether the polynomial is irreducible or not, and finally
for degree 4 and higher, Kronecker's factorization algorithm is applied.

\input{session}



\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
