\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{An Isabelle/HOL formalization of Strong Security}
\author{Sylvia Grewe, Alexander Lux, Heiko Mantel, Jens Sauer}
\maketitle

\begin{abstract}
  Research in information-flow security aims at developing methods to
  identify undesired information leaks within programs from private
  sources to public sinks. Noninterference captures this
  intuition. Strong security from \cite{sabelfeld2000probabilistic}
  formalizes noninterference for concurrent systems.

  We present an Isabelle/HOL formalization of strong security for
  arbitrary security lattices (\cite{sabelfeld2000probabilistic} uses
  a two-element security lattice). The formalization includes
  compositionality proofs for strong security and a soundness proof
  for a security type system that checks strong security for programs
  in a simple while language with dynamic thread creation.

  Our formalization of the security type system is abstract in the
  language for expressions and in the semantic side conditions for
  expressions. It can easily be instantiated with different syntactic
  approximations for these side conditions. The soundness proof of
  such an instantiation boils down to showing that these syntactic
  approximations imply the semantic side conditions.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
%\input{session}

\section{Preliminary definitions}

\subsection{Type synonyms}

The formalization is parametric in different aspects. Notably, it is
parametric in the security lattice it supports. 

For better readability, we use the following type synonyms in our formalization:

\input{Types.tex}

\section{Strong security}

\subsection{Definition of strong security}

We define strong security such that it is parametric in a security
lattice (\textit{'d}). The definition of strong security by itself is
language-independent, therefore the definition is parametric in a
programming language (\textit{'com}) in addition.

\input{Strong_Security.tex}


\subsection{Proof technique for compositionality results}

For proving compositionality results for strong security, we formalize
the following ``up-to technique'' and prove it sound:

\input{Up_To_Technique.tex}


\subsection{Proof of parallel compositionality}

We prove that strong security is preserved under composition of
strongly secure threads.

\input{Parallel_Composition.tex}


\section{Example language and compositionality proofs}

\subsection{Example language with dynamic thread creation}

As in \cite{sabelfeld2000probabilistic}, we instantiate the language
with a simple while language that supports dynamic thread creation via
a fork command (Multi-threaded While Language with fork, MWLf). Note
that the language is still parametric in the language used for Boolean
and arithmetic expressions (\textit{'exp}).

\input{MWLf.tex}

\subsection{Proofs of atomic compositionality results}

We prove for each atomic command of our example programming language
(i.e. a command that is not composed out of other commands) that it is
strongly secure if the expressions involved are indistinguishable for
an observer on security level $d$.

\input{Strongly_Secure_Skip_Assign.tex}

\subsection{Proofs of non-atomic compositionality results}

We prove compositionality results for each non-atomic command of our
example programming language (i.e. a command that is composed out
of other commands): If the components are strongly secure and the
expressions involved indistinguishable for an observer on security
level $d$, then the composed command is also strongly secure.

\input{Language_Composition.tex}

\section{Security type system}

\subsection{Abstract security type system with soundness proof}

We formalize an abstract version of the type system in
\cite{sabelfeld2000probabilistic} using locales
\cite{conf/types/Ballarin03}. Our formalization of the type system is
abstract in the sense that the rules specify abstract semantic side
conditions on the expressions within a command that satisfy for
proving the soundness of the rules. That is, it can be instantiated
with different syntactic approximations for these semantic side
conditions in order to achieve a type system for a concrete language
for Boolean and arithmetic expressions. Obtaining a soundness proof
for such a concrete type system then boils down to proving that the
concrete type system interprets the abstract type system.

We prove the soundness of the abstract type system by simply applying
the compositionality results proven before.

\input{Type_System.tex}

\subsection{Example language for Boolean and arithmetic expressions}

As and example, we provide a simple example language for instantiating the parameter
\textit{'exp} for the language for Boolean and arithmetic expressions.

\input{Expr.tex}

\subsection{Example interpretation of abstract security type system}

Using the example instantiation of the language for Boolean and
arithmetic expressions, we give an example instantiation of our
abstract security type system, instantiating the parameter for domains
\textit{'d} with a two-level security lattice.

\input{Domain_example.tex}

\input{Type_System_example.tex}


% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
