\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}

\renewcommand{\isasymlonglonglongrightarrow}{$\longrightarrow$}


\begin{document}

\title{Matrices for ODEs}
\author{Jonathan Juli\'an Huerta y Munive}
\maketitle

\begin{abstract}
  Our theories formalise various matrix properties that serve to establish 
  existence, uniqueness and characterisation of the solution to affine 
  systems of ordinary differential equations (ODEs). In particular, we
  formalise the operator and maximum norm of matrices. Then we use 
  them to prove that square matrices form a Banach space, and in this 
  setting, we show an instance of Picard-Lindel\"of's theorem for affine
  systems of ODEs. Finally, we apply this formalisation by verifying three 
  simple hybrid programs.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introductory Remarks}

Affine systems of ordinary differential equations (ODEs) are those whose associated vector fields are linear transformations. That is, if there is a matrix-valued function $A:\mathbb{R}\to M_{n\times n}(\mathbb{R})$ and vector function $B:\mathbb{R}\to\mathbb{R}^n$ such that the system of ODEs $x'\, t=f\, (t,x\, t)$ can be rewritten as $x'\, t=A\cdot (x\, t)+B\, t$, then the system is affine. Similarly, the associated linear system of ODEs is $x'\, t=A\cdot (x\, t)$ for matrix-vector multiplication $\cdot$. Our theories formalise affine (hence linear) systems of ordinary differential equations. For this purpose, we extend the ODE libraries of~\cite{ImmlerH12a} and linear algebra in HOL-Analysis. We add to them various results about invertibility of matrices, their diagonalisation, their operator and maximum norms, and properties relating them with vectors. We also define a new type of square matrices and prove that this is a Banach space. Then we obtain results about derivatives of matrix-vector multiplication and use them to prove Picard-Lindel\"of's theorem as formalised in~\cite{afp:hybrid}. The Banach space instance allows us to characterise the general solution to affine systems of ODEs in terms of the matrix-exponential. Finally, we use the components of~\cite{afp:hybrid} to do three simple verification examples in the style of differential dynamic logic~\cite{Platzer10} as showcased in~\cite{ArmstrongGS16,FosterMS19,MuniveS19}. A paper with a detailed overview of the various contributions that this formalisation adds to the verification components will be available soon in ArXiv.

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
