\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage[T1]{fontenc}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\renewcommand{\isadigit}[1]{\ensuremath{#1}}
\protected\def\isacharunderscore{\raisebox{2pt}{\_\kern-1.7pt}}

\begin{document}

\title{Weight-Balanced Trees}
\author{Tobias Nipkow and Stefan Dirix}
\maketitle

\begin{abstract}
  This theory provides a verified implementation of weight-balanced trees
  following the work of Hirai and Yamamoto \cite{HiraiY11} who proved that all
  parameters in a certain range are valid, i.e. guarantee that insertion and
  deletion preserve weight-balance.  Instead of a general theorem we provide
  parameterized proofs of preservation of the invariant that work for many (all?)
  valid parameters.
\end{abstract}

\section{Introduction}

Weight-balanced trees (\emph{WB} trees) are a class of binary search trees of logarithmic height.
They were invented by Nievergelt and Reingold \cite{NievergeltR72,NievergeltR73}
who called them \emph{trees of bounded balance}. They are parametrized by a constant.
Parameters are called \emph{valid} if they guarantee that insertion and deletion preserve
the WB invariant. Blum and Mehlhorn \cite{BlumM80} later discovered that there is a
flaw in Nievergelt and Reingold's analysis of valid parameters
and gave a detailed correctness proof for a modified range of parameters.
Adams \cite{Adams92,Adams93} considered a slightly modified version
of WB trees and analyzed which parameters are valid.
The Haskell libraries \texttt{Data.Set} and \texttt{Data.Map} are based on Adams' papers
but it was found that the implementation did not preserve the invariant. This motivated
Hirai and Yamamoto \cite{HiraiY11} to verify the valid parameter range for the original
definition of WB tree formally in Coq.
They also showed that Adams' analysis is flawed by giving a counterexample to Adams' claimed
range of valid parameters. Straka \cite{Straka12} analyzes valid parameters for Adam's variant.
Yet another variant of WB trees was considered by Roura \cite{Roura01}.

% include generated text of all theories
\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
