
\documentclass[10pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{fullpage}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Declarative Semantics for Functional Languages}
\author{Jeremy G. Siek}
\maketitle

\begin{abstract}
We present a semantics for an applied call-by-value
lambda-calculus that is compositional, extensional, and
elementary. We present four different views of the semantics: 1)
as a relational (big-step) semantics that is not operational but
instead declarative, 2) as a denotational semantics that does not
use domain theory, 3) as a non-deterministic interpreter, and 4)
as a variant of the intersection type systems of the Torino
group.  We prove that the semantics is correct by showing that it
is sound and complete with respect to operational semantics on
programs and that is sound with respect to contextual
equivalence. We have not yet investigated whether it is fully
abstract. We demonstrate that this approach to semantics is
useful with three case studies. First, we use the semantics to
prove correctness of a compiler optimization that inlines
function application. Second, we adapt the semantics to the
polymorphic lambda-calculus extended with general recursion and
prove semantic type soundness.  Third, we adapt the semantics to
the call-by-value lambda-calculus with mutable references.
The paper that accompanies these Isabelle theories is available
on arXiv at the following URL: \\
\url{https://arxiv.org/abs/1707.03762}
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\pagebreak

% generated text of all theories
\input{session}

% optional bibliography
%\bibliographystyle{abbrv}
%\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
