\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage[utf8]{inputenc}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\newtheorem{theorem}{Theorem}%[section]
\newtheorem{corollary}{Corollary}%[section]
\newcommand\isafor{\textsf{IsaFoR}}
\newcommand\ceta{\textsf{Ce\kern-.18emT\kern-.18emA}}
\newcommand\rats{\mathbb{Q}}
\newcommand\ints{\mathbb{Z}}
\newcommand\reals{\mathbb{R}}
\newcommand\complex{\mathbb{C}}

\newcommand\rai{real algebraic number\xspace}
\newcommand\rais{real algebraic numbers\xspace}

\begin{document}

\title{Perron-Frobenius Theorem for Spectral Radius Analysis\footnote{Supported by FWF (Austrian Science Fund) project Y757.}}
\author{Jose Divas\'on, Ondřej Kunčar, Ren\'e Thiemann and Akihisa Yamada}
\maketitle

\begin{abstract}
The spectral radius of a matrix $A$ is the maximum
norm of all eigenvalues of $A$. In previous work we already formalized that 
for a complex matrix $A$, the values in $A^n$ grow polynomially in $n$ if and only if the spectral radius is 
at most one.
One problem with the above characterization is the determination of all \emph{complex} eigenvalues.
In case $A$ contains only non-negative real values, a simplification is possible with the help of
the Perron-Frobenius theorem, which tells us that 
it suffices to consider only the \emph{real} eigenvalues of $A$, i.e., 
applying Sturm's method can decide the polynomial growth of $A^n$.

We formalize the Perron-Frobenius theorem based on a proof via Brouwer's fixpoint theorem, which
is available in the HOL multivariate analysis (HMA) library. Since the results on the spectral
radius is based on matrices in the Jordan normal form (JNF) library, we further develop
a connection which allows us to easily transfer theorems between HMA and JNF. 
With this connection we derive the combined result: if $A$ is a non-negative real matrix,
and no real eigenvalue of $A$ is strictly larger than one, then $A^n$ is polynomially bounded in $n$.
\end{abstract}

\tableofcontents

\section{Introduction}

The spectral radius of a matrix $A$ over $\reals$ or $\complex$ 
is defined as \begin{equation*}
\rho(A) = \max\,\{|x| .\ \chi_A(x) = 0, x \in \complex\}
\end{equation*} 
where
$\chi_A$ is the characteristic polynomial of $A$. It is a central notion related to the growth rate
of matrix powers. A matrix $A$ has polynomial growth, i.e., 
all values of $A^n$ can be bounded polynomially in $n$, if and only if $\rho(A) \leq 1$.
It is quite easy to see that $\rho(A) \leq 1$ is a  necessary criterion,\footnote{
Let $\lambda$ and $v$ be some eigenvalue and eigenvector pair such that $|\lambda| > 1$. Then
$|A^n v| = |\lambda^n v| = |\lambda|^n |v|$ grows exponentially in $n$, where $|w|$ denotes the component-wise
application of $|\cdot|$ to vector elements of $w$.}
but it is more complicated to argue about sufficiency.
In previous work we formalized this statement via Jordan normal forms
\cite{JNF}.

\begin{theorem}[in JNF]
\label{sr}
The values in $A^n$ are polynomially bounded in $n$ if $\rho(A) \leq 1$.
\end{theorem}

In order to perform the proof via Jordan normal forms, we did not use the HMA library from the distribution
to represent matrices. The reason is that already the definition of a Jordan normal form is naturally
expressed via block-matrices, and arbitrary block-matrices are hard to express in HMA, if at all.

The problem in applying Theorem~\ref{sr} in concrete examples is the determination of all
complex roots of the polynomial $\chi_A$. For instance, one can utilize complex algebraic numbers 
for this purpose, which however are computationally expensive. To avoid this problem, in this work we 
formalize the Perron Frobenius theorem. It states that for non-negative real-valued matrices,
$\rho(A)$ is an eigenvalue of $A$.

\begin{theorem}[in HMA]
\label{pf}
If $A \in \reals_{\geq 0}^{k \times k}$, then $\chi_A(\rho(A)) = 0$.
\end{theorem}

We decided to perform the formalization based on the HMA library, since there is a short proof
of Theorem~\ref{pf} via Brouwer's fixpoint theorem \cite[Section 5.2]{SerreMatrices}. The latter is a well-known but complex theorem 
that is available in HMA, but not in the JNF library.

Eventually we want to combine both theorems to obtain:
\begin{corollary}
\label{final}
If $A \in \reals_{\geq 0}^{k \times k}$, then the values in $A^n$ are polynomially bounded in
$n$ if $\chi_A$ has no real roots in the interval $(1,\infty)$. 
\end{corollary}

This criterion is computationally far less expensive -- one invocation of Sturm's method 
on $\chi_A$ suffices. Unfortunately, we cannot immediately combine both theorems. We first have to bridge
the gap between the HMA-world and the JNF-world. To this end, we develop a setup for the transfer-tool
which admits to translate theorems from JNF into HMA. Moreover, using a recent extension for local type
definitions within proofs \cite{LTD}, we also provide a translation from HMA into JNF. 


With the help of these translations, we prove Corollary~\ref{final} and make it available in
both HMA and JNF. (In the formalization the corollary looks a bit more complicated as it also
contains an estimation of the the degree of the polynomial growth.)

% include generated text of all theories
\input{session}

\paragraph*{Acknowledgements}
We thank Fabian Immler for an introduction to continuity proving using HMA.

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
