\documentclass[12pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% additional symbol fonts
\usepackage{stmaryrd}
\usepackage{amssymb}

% this should be the last package used
\usepackage{pdfsetup}

\usepackage{times}
%\usepackage{a4med}

\newcommand{\isasymiinter}{\isamath{\doublecap}}
\newcommand{\isasymocirc}{\isamath{\circledcirc}}
\newcommand{\isasymostar}{\isamath{\varoast}}

\newcommand{\quotient}{\mathbin{//}}
\newcommand{\Seq}{\mathbin{;}}
\newcommand{\refsto}{\mathrel{\sqsubseteq}}
\newcommand{\nondet}{\mathbin{\sqcap}}
\newcommand{\Nondet}{\mathop{\bigsqcap}}
\newcommand{\together}{\mathbin{\doublecap}}
%\newcommand{\parallel}{\mathbin{||}}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{Concurrent Refinement Algebra and Rely Quotients}
\author{Julian Fell and Ian Hayes and Andrius Velykis}
\maketitle

\begin{abstract}
The concurrent refinement algebra developed here is designed to provide a 
foundation for rely/guarantee reasoning about concurrent programs.
The algebra builds on a complete lattice of commands by providing
sequential composition, parallel composition and a novel weak conjunction operator.
The weak conjunction operator coincides with the lattice supremum 
providing its arguments are non-aborting, but aborts if either of its arguments do.
Weak conjunction provides an abstract version of a guarantee condition
as a guarantee process.
We distinguish between models that distribute sequential composition
over non-deterministic choice from the left 
(referred to as being conjunctive in the refinement calculus literature)
and those that don't.
Least and greatest fixed points of monotone functions are provided to
allow recursion and iteration operators to be added to the language.
Additional iteration laws are available for conjunctive models.
The rely quotient of processes $c$ and $i$ is the process that, 
if executed in parallel with $i$ implements $c$.
It represents an abstract version of a rely condition generalised to a process.
\end{abstract}

\newpage
\tableofcontents
\parindent 0pt\parskip 0.5ex

\newpage
\section{Overview}

The theories provided here were developed in order to provide support 
for rely/guarantee concurrency \cite{Jones81d,jon83a}.
The theories provide a quite general concurrent refinement algebra
that builds on a complete lattice of commands
by adding sequential and parallel composition operators
as well as recursion.
A novel weak conjunction operator is also added as this allows one to build
more general specifications.
The theories are based on the paper by Hayes~\cite{AFfGRGRACP},
however there are some differences that have been introduced to correct and simplify
the algebra and make it more widely applicable.
See the appendix for a summary of the differences.

The basis of the algebra is a complete lattice of commands (Section~\ref{S:lattice}).
Sections~\ref{S:sequential}, \ref{S:parallel} and \ref{S:conjunction} 
develop laws for sequential composition, parallel composition and weak conjunction,
respectively, based on the refinement lattice.
Section~\ref{S:CRA} brings the above theories together.
Section~\ref{S:galois} adds least and greatest fixed points and there associated laws,
which allows finite, possibly infinite and strictly infinite iteration operators to
be defined in Section~\ref{S:iteration} in terms of fixed points.

The above theories do not assume that sequential composition is conjunctive.
Section~\ref{S:conjunctive-sequential} adds this assumption and derives a further
set of laws for sequential composition and iterations.

Section~\ref{S:rely-quotient} builds on the general theory to provide a rely quotient
operator that can be used to provide a general rely/guarantee framework for 
reasoning about concurrent programs.

\input{session}

\section{Conclusions}

The theories presented here provide a quite abstract view of the rely/guarantee approach
to concurrent program refinement.
A trace semantics for this theory has been developed \cite{DaSMfaWSLwC}.
The concurrent refinement algebra is general enough to also form the basis of a
more concrete rely/guarantee approach based on a theory of atomic steps and
synchronous parallel and weak conjunction operators \cite{FM2016atomicSteps}.

\subparagraph*{Acknowledgements.}

This research was supported by
Australian Research Council Grant grant
DP130102901
and
EPSRC (UK) Taming Concurrency grant.
This research has benefited from feedback from
Robert Colvin,
Chelsea Edmonds,
Ned Hoy,
Cliff Jones,
Larissa Meinicke,
and
Kirsten Winter.
%but the remaining errors are all courtesy of the authors.

\appendix
\section{Differences to earlier paper}

This appendix summarises the differences between these Isabelle theories and
the earlier paper \cite{AFfGRGRACP}.
We list the changes to the axioms but not all the flow on effects to lemmas.

\begin{enumerate}
\item
The earlier paper assumes $c \Seq (d_0 \nondet d_1) = (c \Seq d_0) \nondet (c \Seq d_1)$
but here we separate the case where this is only a refinement from left to right (Section~\ref{S:sequential})
from the equality case (Section~\ref{S:conjunctive-sequential}).
\item\label{diff:distr-par}
The earlier paper assumes $(\Nondet C) \parallel d = (\Nondet c \in C . c \parallel d)$
but in Section~\ref{S:parallel}  we assume this only for non-empty $C$ and furthermore assume 
that parallel is abort strict, i.e. $\bot \parallel c = c$.
\item
The earlier paper assumes $c \together (\bigsqcup D) = (\bigsqcup d \in D . c \together d)$.
In Section~\ref{S:conjunction} that assumption is not made
because it does not hold for the model we have in mind \cite{DaSMfaWSLwC}
but we do assume $c \together \bot = \bot$.
\item
In Section~\ref{S:CRA} we add the assumption $nil \refsto nil \parallel nil$ to locale sequential-parallel.
\item
In Section~\ref{S:CRA} we add the assumption $\top \refsto chaos \parallel \top$.
\item
In Section~\ref{S:CRA} we assume only $chaos \refsto chaos \parallel chaos$
whereas in the paper this is an equality (the reverse direction is straightforward to prove).
\item
In Section~\ref{S:CRA} axiom chaos-skip ($chaos \refsto skip$) has been dropped
because it can be proven as a lemma using the parallel-interchange axiom.
\item
In Section~\ref{S:CRA} we add the assumption $chaos \refsto chaos \Seq chaos$.
\item
Section~\ref{S:conjunctive-sequential} assumes
$D \neq \{\} \Rightarrow c \Seq \Nondet D = (\Nondet d \in D . c \Seq d)$.
This distribution axiom is not considered in the earlier paper.
\item
Because here parallel does not distribute over an empty non-deterministic choice
(see point \ref{diff:distr-par} above)
in Section~\ref{S:rely-quotient} the theorem rely-quotient needs to assume the interference process $i$
is non-aborting (refines chaos). This also affects many lemmas in this section that depend on 
theorem rely-quotient.
\end{enumerate}


\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
