\documentclass[11pt,a4paper, DIV=11]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fullpage}
\usepackage{isabelle,isabellesym}
\usepackage{amssymb}
\usepackage{amsmath,amsfonts}

\renewcommand{\bf}{\normalfont\bfseries}
\renewcommand{\rm}{\normalfont\rmfamily}
\renewcommand{\it}{\normalfont\itshape}

\usepackage{pdfsetup}

\hypersetup{
  pdfinfo={
    Title={The IMAP CmRDT},
    Subject={},
    Keywords={IMAP, Isabelle, CRDT},
    Author={Tim Jungnickel, Lennart Oldenburg, Matthias Loibl},
    Creator={}
  },
  bookmarksopen=true,
  bookmarksnumbered,
  bookmarksopenlevel=2,
  bookmarksdepth=3
}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\renewcommand{\isastyle}{\isastyleminor}

\usepackage{algorithm} % for CRDT translation
\usepackage[noend]{algpseudocode} % for CRDT translation
\usepackage{amssymb} % for correct math symbols

\newcommand{\create}{\textit{create}}
\newcommand{\delete}{\textit{delete}}
\newcommand{\store}{\textit{store}}
\newcommand{\append}{\textit{append}}
\newcommand{\expunge}{\textit{expunge}}

\newcommand{\session}{\textit{session}}


\title{The IMAP CmRDT}
\author{Tim Jungnickel, Lennart Oldenburg, Matthias Loibl}
\date{\today}


\begin{document}


\maketitle


\begin{abstract}
We provide our Isabelle/HOL formalization of a Conflict-free Replicated Data Type
for Internet Message Access Protocol commands. To this end, we show that Strong
Eventual Consistency (SEC) is guaranteed by proving the commutativity of concurrent
operations. We base our formalization on the recently proposed "framework for
establishing Strong Eventual Consistency for Conflict-free Replicated Datatypes"
(AFP.CRDT) by Gomes et al{.} Hence, we provide an additional example of how the
recently proposed framework can be used to design and prove CRDTs.
\end{abstract}


\tableofcontents
%\newpage


\section{Preface}

A Conflict-free Replicated Data Type (CRDT) \cite{shapiro_crdt} ensures
convergence of replicas without requiring a central coordination server or
even a distributed coordination system based on consensus or locking.
Despite the fact that Shapiro et al{.} provide a comprehensive collection
of definitions for the most useful data types such as registers, sets, and
lists \cite{shapiro_report}, we observe that the use of CRDTs in standard IT
services is rather uncommon. Therefore, we use the Internet Message Access
Protocol (IMAP)---the de-facto standard protocol to retrieve and manipulate mail
messages on an email server---as an example to show the feasibility of using
CRDTs for replicating state of a standard IT service to achieve planetary scale.

Designing a \emph{correct} CRDT is a challenging task. A CmRDT, the
operation-based variant of a CRDT, requires all operations to commute.
To this end, Gomes et al{.} recently published a CmRDT verification
framework \cite{gomes_crdtafp} in Isabelle/HOL.

In our most recent work \cite{pluto}, we presented \emph{pluto}, our research
prototype of a planetary-scale IMAP service. To achieve the claimed
planet-scale, we designed a CmRDT that provides multi-leader replication
of mailboxes without the need of synchronous operations. In order to ensure
the correctness of our proposed IMAP CmRDT, we implemented it in the
verification framework proposed by Gomes et al{.}

In this work, we present our Isabelle/HOL proof of the necessary properties and
show that our CmRDT indeed guarantees Strong Eventual Consistency (SEC).
We contribute not only the certainty that our CmRDT design is correct,
but also provide one more example of how the verification framework
can be used to prove the correctness of a CRDT.


\subsection{The IMAP CmRDT}

In the rest of this work, we show how we modeled our IMAP CmRDT in Isabelle/HOL.
We start by presenting the original IMAP CmRDT, followed by the implementation
details of the Isabelle/HOL formalization. The presentation of our CmRDT in
Spec.~\ref{spec:imap} is based on the syntax introduced in \cite{shapiro_report}.
We highly recommend reading the foundational work by Shapiro et al{.} prior to
following our proof documentation.

In essence, the IMAP CmRDT represents the state of a mailbox, containing folders
(of type $\mathcal{N}$) and messages (of type $\mathcal{M}$). Moreover, we introduce
metadata in form of tags (of type $\texttt{ID}$). All modeling details and a more
detailed description of the CmRDT are provided in the original paper \cite{pluto}.

\begin{algorithm}[t]
  \floatname{algorithm}{Specification}
  \caption{The IMAP CmRDT} \label{spec:imap}
  \algsetblock{payload}{}{1}{0.5cm}
  \algsetblock{update}{}{2}{0.5cm}

  \algsetblockdefx{atsourceone}{}{1}{0.5cm}[1][]{\textbf{atSource} #1}{}
  \algsetblockdefx{atsourcetwo}{}{2}{0.5cm}[1][]{\textbf{atSource} #1}{}
  \algsetblockdefx{downstreamone}{}{1}{0.5cm}[1][]{\textbf{downstream} #1}{}

  \begin{algorithmic}[1]
    \payload \ map $u: \mathcal{N} \rightarrow \mathcal{P}(\texttt{ID}) \times
    \mathcal{P}(\mathcal{M})$ \Comment{$\{\text{foldername}\ f
    \mapsto (\{\text{tag}\ t\}, \{\text{msg}\ m\}), \dots \}$}
      \State initial $\left(\lambda x . (\varnothing, \varnothing)\right)$
      \vspace{0.3em}
    \update \ \create\ $(\text{foldername}\ f)$
      \atsourceone{}
        \State let $\alpha = \textit{unique}()$
      \downstreamone{$(f, \alpha)$}
        \State $u(f) \mapsto (u(f)_1 \cup \{\alpha\}, u(f)_2)$ \vspace{0.3em}
    \update \ \delete\ $(\text{foldername}\ f)$
      \atsourcetwo{$(f)$}
        \State let $R_1 = u(f)_1$
        \State let $R_2 = u(f)_2$
      \downstreamone{$(f, R_1, R_2)$}
        \State $u(f) \mapsto (u(f)_1 \setminus R_1, u(f)_2 \setminus R_2)$
        \update \ \append\ $(\text{foldername}\ f, \text{message}\ m)$
    \atsourceone{$(m)$}
      \State \textbf{pre} $m$ is globally unique
    \downstreamone{$(f, m)$}
      \State $u(f) \mapsto (u(f)_1, u(f)_2 \cup \{m\})$ \vspace{0.3em}
    \update \ \expunge\ $(\text{foldername}\ f, \text{message}\ m)$
      \atsourcetwo{$(f, m)$}
        \State \textbf{pre} $m \in u(f)_2$
        \State let $\alpha = \textit{unique}()$
      \downstreamone{$(f, m, \alpha)$}
        \State $u(f) \mapsto (u(f)_1 \cup \{\alpha\}, u(f)_2 \setminus \{m\})$ \vspace{0.3em}
    \update \ \store\ $(\text{foldername}\ f, \text{message}\ m_\text{old},
    \text{message}\ m_\text{new})$
      \atsourcetwo{$(f, m_\textit{old}, m_\textit{new})$}
        \State \textbf{pre} $m_\text{old} \in u(f)_2$
        \State \textbf{pre} $m_\text{new}$ is globally unique
      \downstreamone{$(f, m_\text{old}, m_\text{new})$}
        \State $u(f) \mapsto (u(f)_1, (u(f)_2 \setminus \{m_\text{old}\}) \cup
      \{m_\text{new}\})$
\end{algorithmic}
\end{algorithm}

The only notable difference between the presented specification and our Isabelle/HOL
formalization is, that we no longer distinguish between sets $\texttt{ID}$ and
$\mathcal{M}$ and that the generated tags of \create\ and \expunge\ are handled
explicitly. This makes the formalization slightly easier, because less type variables
are introduced. The concrete definition can be found in the \textit{IMAP-CRDT Definitions}
section of the \texttt{IMAP-def.thy} file.


\subsection{Proof Guide}

\textit{Hint:} In our proof, we build on top of the definitions given by Gomes et al{.}
in \cite{gomes_crdtisabelle}. We strongly recommend to read their paper first before
following our proof. In fact, in our formalization we reuse the \textit{locales} of the
proposed framework and therefore this work cannot be compiled without the reference
to \cite{gomes_crdtafp}.

Operation-based CRDTs require all concurrent operations to commute in order to ensure
convergence. Therefore, we begin our verification by proving the commutativity of every
combination of possible concurrent operations. Initially, we used \textit{nitpick} to
identify corner cases in our implementation. We prove the commutativity in Section 3
of the \texttt{IMAP-proof-commute.thy} file. The \textit{critical conditions} to satisfy
in order to commute, can be summarized as follows:
\begin{itemize}
  \item The tags of a \create\ and \expunge\ operation or the messages of an
  \append\ and \store\ operation are never in the removed-set of a concurrent
  \delete\ operation.
  \item The message of an \append\ operation is never the message that is
  deleted by a concurrent \store\ or \expunge\ operation.
  \item The message inserted by a \store\ operation is never the message
  that is deleted by a concurrent \store\ or \expunge\ operation.
\end{itemize}

The identified conditions obviously hold in regular traces of our system, because
an item that has been inserted by one operation cannot be deleted by a concurrent
operation. It simply cannot be present at the time of the initiation of the
concurrent operation.

Next, we show that the identified conditions actually hold for all concurrent
operations. Because all tags and all inserted messages are globally unique, it can
easily be shown that all conditions are satisfied. In Isabelle/HOL, showing this fact
takes some effort. Fortunately, we were able to reuse parts of the Isabelle/HOL implementation
of the OR-Set proof in \cite{gomes_crdtafp}. The Isabelle/HOL proofs for the
\textit{critical conditions} are encapsulated in the \texttt{IMAP-proof-independent.thy} file.

With the introduced lemmas, we prove the final theorem that states that convergence
is guaranteed. Due to all operations being commutative in case the \textit{critical conditions}
are satisfied and the \textit{critical conditions} indeed are holding for all concurrent updates,
all concurrent operations commute. The Isabelle/HOL proof is contained in the
\texttt{IMAP-proof.thy} file.

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
