\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{wasysym}%Needed for next (circle)
\usepackage{fullpage}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}

\begin{document}

\title{Dynamic Architectures%
%\thanks{The final publication is available at the Archive of Formal Proofs via \url{http://isa-afp.org/entries/DynamicArchitectures.shtml}.}%
}
\author{Diego Marmsoler}
\maketitle

\begin{abstract}
	%Context
	The architecture of a system describes the system's overall organization into components and connections between those components.
	With the emergence of mobile computing, dynamic architectures have become increasingly important.
	In such architectures, components may appear or disappear, and connections may change over time.
	%Problem
	In the following we mechanize a theory of dynamic architectures and verify the soundness of a corresponding calculus.
	%Approach
	Therefore, we first formalize the notion of configuration traces~\cite{Marmsoler2016} as a model for dynamic architectures.
	Then, the behavior of single components is formalized in terms of behavior traces and an operator is introduced and studied to extract the behavior of a single component out of a given configuration trace.
	Then, behavior trace assertions are introduced as a temporal specification technique to specify behavior of components.
	Reasoning about component behavior in a dynamic context is formalized in terms of a calculus for dynamic architectures~\cite{Marmsoler2017c}.
	Finally, the soundness of the calculus is verified by introducing an alternative interpretation for behavior trace assertions over configuration traces and proving the rules of the calculus.
	Since projection may lead to finite as well as infinite behavior traces, they are formalized in terms of coinductive lists.
	Thus, our theory is based on Lochbihler's~\cite{Lochbihler2010} formalization of coinductive lists.
	%Implications
	The theory may be applied to verify properties for dynamic architectures.
\end{abstract}
\newpage

\tableofcontents
\newpage

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
