\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym,latexsym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Signature-Based Gr\"obner Basis Algorithms}
\author{Alexander Maletzky\thanks{Supported by the Austrian Science Fund (FWF): P 29498-N31}}
\maketitle

\begin{abstract}
This article formalizes signature-based algorithms for computing Gr\"obner bases. Such 
algorithms are, in general, superior to other algorithms in terms of efficiency, and have not been 
formalized in any proof assistant so far. The present development is both generic, in the sense that 
most known variants of signature-based algorithms are covered by it, and effectively executable on 
concrete input thanks to Isabelle's code generator. Sample computations of benchmark problems show 
that the verified implementation of signature-based algorithms indeed outperforms the existing 
implementation of Buchberger's algorithm in Isabelle/HOL.

Besides total correctness of the algorithms, the article also proves that under certain 
conditions they a-priori detect and avoid all useless zero-reductions, and always return `minimal' 
(in some sense) Gr\"obner bases if an input parameter is chosen in the right way.

The formalization follows the recent survey article by Eder and Faug\`ere.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\newpage
\section{Introduction}

Signature-based algorithms~\cite{Faugere2002,Eder2017} play are central role in modern computer 
algebra systems, as they allow to compute Gr\"obner bases of ideals of multivariate polynomials much 
more efficiently than other algorithms. Although they also belong to the class of 
critical-pair/completion algorithms, as almost all algorithms for computing Gr\"obner bases, they 
nevertheless possess some quite unique features that render a formal development in proof assistants 
challenging. In fact, this is the first formalization of signature-based algorithms in any proof 
assistant.

The formalization builds upon the existing formalization of Gr\"obner bases 
theory~\cite{Immler2016} and closely follows Sections~4--7 of the excellent survey 
article~\cite{Eder2017}. Some proofs were taken from~\cite{Roune2012,Eder2013}.

Summarizing, the main features of the formalization are as follows:
\begin{itemize}
  \item It is \emph{generic}, in the sense that it considers the computation of so-called 
\emph{rewrite bases} and neither fixes the term order nor the rewrite-order.

  \item It is \emph{efficient}, in the sense that all executable algorithms (e.\,g. 
\textit{gb-sig}) operate on sig-poly-pairs rather than module elements, and that polynomials are 
represented efficiently using ordered associative lists.

  \item It proves that if the input is a regular sequence and the term order is a POT order, there
are no useless zero-reductions (Theorem \textit{gb-sig-no-zero-red}).

  \item It proves that the signature Gr\"obner bases computed w.\,r.\,t. the `ratio' rewrite order 
are minimal (Theorem \textit{gb-sig-z-is-min-sig-GB}).

  \item It features sample computations of benchmark problems to illustrate the practical usability 
of the verified algorithms.
\end{itemize}

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
