\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{amssymb}
% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\newcommand\isafor{\textsf{IsaFoR}}
\newcommand\ceta{\textsf{Ce\kern-.18emT\kern-.18emA}}
\newcommand\nats{\mathbb{N}}
\newcommand\reals{\mathbb{R}}
\newcommand\rats{\mathbb{Q}}
\newcommand\fieldext[2]{#1[#2]}
\newcommand\ratsb{\fieldext\rats{\sqrt b}}

\begin{document}

\title{Implementing field extensions of the form $\ratsb$\thanks{This research is supported by FWF (Austrian Science Fund) project P22767-N13.}}
\author{Ren\'e Thiemann}
\maketitle

\begin{abstract}
  We apply data refinement to implement the real numbers, where we support
  all numbers in the field extension $\ratsb$, i.e., all numbers of the 
  form $p + q \sqrt{b}$ for rational numbers $p$ and $q$ and some fixed natural
  number $b$. To this end, we also developed algorithms to precisely compute
  roots of a rational number, and to perform a factorization of natural 
  numbers which eliminates duplicate prime factors.
  
  Our results have been used to certify termination proofs which involve
  polynomial interpretations over the reals.  
\end{abstract}



\tableofcontents

\section{Introduction}

It has been shown that polynomial
interpretations over the reals are strictly more powerful for termination
proving than polynomial interpretations over the rationals. To this end,
also automated termination prover started to generate such interpretations.
\cite{Rational,Luc06,Luc07,LPAR09,SCSS10}. However, for all current implementations,
only reals of the form $p + q \cdot \sqrt{b}$ are generated where $b$ is some fixed
natural number and $p$ and $q$ may be arbitrary rationals, i.e., we get
numbers within $\ratsb$.

To support these termination proofs in our certifier \ceta\ \cite{CeTA}, 
we therefore required executable functions on $\ratsb$, which can then
be used as an implementation type for the reals. Here, we used ideas from
\cite{datarefinement,Loc13} to provide a sufficiently powerful 
partial implementations via data refinement.


\input{session}

\section*{Acknowledgements}
We thank Bertram Felgenhauer for interesting discussions and especially
for mentioning Cauchy's mean theorem during the formalization of
the algorithms for computing roots.

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
