\documentclass[11pt,a4paper]{report}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

%\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{CakeML}
\author{Lars Hupel, Yu Zhang}
\maketitle

\begin{abstract}
  CakeML is a functional programming language with a proven-correct compiler and runtime system.
  This entry contains an unofficial version of the CakeML semantics that has been exported from the Lem specifications to Isabelle.
  Additionally, there are some hand-written theory files that adapt the exported code to Isabelle and port proofs from the HOL4 formalization, e.g.\ termination and equivalence proofs.
\end{abstract}

\tableofcontents

\clearpage

\section*{Contributors}

The export script has been written by Lars Hupel.
Hand-written theory files, including definitions and proofs, have been developed by Lars Hupel and Yu Zhang.

Lem is a project by Peter Sewell et.al.
Contributors can be found on its project page\footnote{\url{https://www.cl.cam.ac.uk/~pes20/lem/}} and on GitHub.\footnote{\url{https://github.com/rems-project/lem/graphs/contributors}}

CakeML is a project with many developers and contributors that can be found on its project page\footnote{\url{https://cakeml.org/}} and on GitHub.\footnote{\url{https://github.com/CakeML/cakeml/graphs/contributors}}

In particular, Fabian Immler and Johannes \AA{}man Pohjola have contributed Isabelle mappings for constants in the Lem specification of the CakeML semantics.

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
%\bibliographystyle{abbrv}
%\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
