\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{shadings,intersections}
% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% notation and presentation style for geometric expressions
\newcommand{\length}[1]{\overline{#1}}

\begin{document}

\title{Intersecting Chords Theorem}
\author{Lukas Bulwahn}
\maketitle

\begin{abstract}

This entry provides a geometric proof of the intersecting chords theorem.
The theorem states that when two chords intersect each other inside a circle,
the products of their segments are equal.

After a short review of existing proofs in the
literature~\cite{proofwiki:chord-theorem, knorr-1989, birkhoff-beatley-1959, harrison},
I decided to use a proof approach that employs reasoning about
lengths of line segments, the orthogonality of two lines and
Pythagoras Law. Hence, one can understand the formalized
proof easily with the knowledge of a few general geometric facts
that are commonly taught in high-school.

This theorem is the 55th theorem of the Top 100 Theorems list.
\end{abstract}

\tableofcontents

\section{Introduction}
The intersecting chords theorem states:
\begin{quote}
When two chords intersect each other inside a circle, the products of their segments are equal.
\end{quote}
To prove this theorem in Isabelle, I reviewed existing formalizations in theorem provers
and proofs in the literature~\cite{proofwiki:chord-theorem, knorr-1989, birkhoff-beatley-1959, harrison}.
At the time of this AFP submission, the formalization of geometry in Isabelle
is limited to only a few concepts and theorems. Hence, I selected to formalize
the proof approach that fitted best to the already existing geometry formalizations.

The proof in HOL Light~\cite{harrison} simply unfolds the involved geometric predicates and then
proves the theorem using only algebraic computations on real numbers. By a quick and
shallow inspection of the proof script without executing the proof script step by step in
HOL Light, I could not understand the proof script well enough to re-write the proof
in Isabelle. As running the script in HOL Light seemed too involved to me, I
ignored HOL Light's proof approach and considered the other approaches in the literature.

The first proof approach~\cite{proofwiki:chord-theorem} that I found in the
literature employs similarity of triangles, the inscribed angle theorem,
and basic reasoning with angles.
The intersecting chords theorem only consists of two reasoning steps after
stating the geometric observations about angles.
However, the proof requires to formalize the concept of similarity of triangles,
extend the existing formalization of angles, and prove the inscribed angle theorem.
So, I abandoned this proof approach and considered the second proof
approach.

The second proof approach~\cite{proofwiki:chord-theorem} needs only
basic geometric reasoning about lengths of line segments, the orthogonality
of two lines and Pythagoras Law.
More specifically, one must prove that the line that goes through
the apex and the midpoint of the base in an isosceles triangle
is orthogonal to the base. This is easily derived from the property
of an isosceles triangle using the congruence properties of triangles,
which is already formalized in AFP's Triangle entry~\cite{Triangle-AFP}.
Furthermore, Pythagoras Law is a special case of the
Law of Cosines, which is already formalized in AFP's Triangle entry.

Ultimately, I decided to use this second proof approach, which I
sketch in more detail in the next subsection.

\subsection{Informal Proof Sketch}

The proof of the intersecting chords theorem relies on the following observation
which is depicted in Figure \ref{fig:chord-property}.

\begin{figure}
\begin{tikzpicture}

\coordinate [label=above:$C$] (C) at (0,0);
\coordinate [label=below:$S$] (S) at (-4,-3);
\coordinate [label=below:$T$] (T) at (4,-3);
\coordinate [label=below:$X$] (X) at (1,-3);
\coordinate [label=below:$M$] (M) at (0,-3);

\draw (S) -- (T);
\draw (S) -- (C);
\draw (T) -- (C);
\draw (M) -- (C);
\draw (X) -- (C);

% draw an arc from S T with center C
% and extend this arc by 10 degrees beyond S and T
% to show a nicely clipped part of the relevant arc.
\pgfmathparse{atan2(-3, -4) - 10};
\pgfmathsetmacro{\startangle}{\pgfmathresult)};
\pgfmathparse{atan2(-3, 4) + 10};
\pgfmathsetmacro{\endangle}{\pgfmathresult)};
\coordinate (S') at (\startangle : 5);
\draw [dashed] (S') arc ( \startangle : \endangle : 5);

\draw (0.2,-3) arc (0:90:0.2);
\coordinate [label=right:\tiny{90\textdegree}] (C) at (0.15,-2.8);
\end{tikzpicture}
\caption{Key Lemma states $\length{SX} \cdot \length{XT} = \length{SC} ^ 2 - \length{XC} ^ 2$}
\label{fig:chord-property}
\end{figure}

Instead of considering \emph{two} arbitrary chords intersecting, consider 
\emph{one} arbitrary chord with endpoints $S$ and $T$ on a circle with center $C$
and one arbitrary point $X$ on the chord $ST$.
This point $X$ on the chord creates two line segments on this chord, the left part $SX$, and
the right part $XT$. Without loss of generality, we can assume that $SX$ is longer that $XT$,
as shown in Figure \ref{fig:chord-property}.

The key lemma for the intersecting chords theorem provides 
a closed expression for the length of these two line segment using
the distances of the chord endpoints and the point to the center $C$, i.e.,
the lemma states:
\begin{quote}
$\length{SX} \cdot \length{XT} = \length{SC} ^ 2 - \length{XC} ^ 2$.
\end{quote}
To prove this fact, we consider the midpoint $M$ of the chord $ST$.
First, as $M$ is the midpoint, $\length{SM}$ and $\length{TM}$ are equal.
Second, we observe that the lengths of the line segments $SX$ and $XT$
are: 
\begin{quote}
$\length{SX} = \length{SM} + \length{MX}$ and
$\length{XT} = \length{TM} - \length{MX} = \length{SM} - \length{MX}$.
\end{quote}
Third, the Pythagoras law for the triangles $SMC$ and $XMC$ states:
\begin{quote}
$\length{SM} ^ 2 + \length{MC} ^ 2 = \length{SC} ^ 2$ and
$\length{XM} ^ 2 + \length{MC} ^ 2 = \length{XC} ^ 2$.
\end{quote}
Finally, the product can be expressed as:
\begin{quote}
$\length{SX} \cdot \length{XT} = (\length{SM} + \length{MX}) \cdot (\length{TM} - \length{MX}) =
\length{SM} ^ 2 - \length{MX} ^ 2 = (\length{SC} ^ 2 - \length{MC} ^ 2) - (\length{XC} ^ 2 - \length{MC} ^ 2) =
\length{SC} ^ 2 - \length{XC} ^ 2$.
\end{quote}
The intersecting chord theorem now follows directly
from this lemma:
as the distances $SC$ and $XC$ for two arbitrary chords
intersecting at $X$ are equal, also the products
of the chord segments are equal.

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

\nocite{*}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
