\documentclass[11pt,a4paper]{article}
\usepackage{textcomp}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath,amssymb}

\usepackage{pdfsetup}

\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{Applicative Lifting}
\author{Andreas Lochbihler \and Joshua Schneider}
\maketitle

\begin{abstract}
  Applicative functors augment computations with effects by lifting function application to types
  which model the effects \cite{mcbride08}.
  As the structure of the computation cannot depend on the effects, applicative expressions can be
  analysed statically.
  This allows us to lift universally quantified equations to the effectful types, as observed by 
  Hinze \cite{hinze10}.
  Thus, equational reasoning over effectful computations can be reduced to pure types.

  This entry provides a package for registering applicative functors and two proof methods for
  lifting of equations over applicative functors.
  The first method applicative{\isacharunderscore}nf normalises applicative expressions according to
  the laws of applicative functors.
  This way, equations whose two sides contain the same list of variables can be lifted to every
  applicative functor.

  To lift larger classes of equations, the second method applicative{\isacharunderscore}\linebreak lifting
  exploits a number of additional properties (e.g., commutativity of effects) provided the
  properties have been declared for the concrete applicative functor at hand upon registration.

  We declare several types from the Isabelle library as applicative functors and illustrate the use of
  the methods with two examples: the lifting of the arithmetic type class hierarchy to streams and
  the verification of a relabelling function on binary trees.
  We also formalise and verify the normalisation algorithm used by the first proof method,
  as well as the general approach of the second method, which is based on bracket abstraction.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
