\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{amsmath} 

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Verification Components for Hybrid Systems}
\author{Jonathan Juli\'an Huerta y Munive}
\maketitle

\begin{abstract}
  These components formalise a semantic framework for the deductive 
  verification of hybrid systems. They support reasoning about continuous 
  evolutions of hybrid programs in the style of differential dynamic logic.
  Vector fields or flows model these evolutions, and their verification is 
  done with invariants for the former or orbits for the latter. Laws of
  modal Kleene algebra or categorical predicate transformers implement 
  the verification condition generation. Examples show the approach at work.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introductory Remarks}

These theories implement verification components for hybrid programs in the style of differential dynamic logic~\cite{Platzer10}, an approach for deductive verification of hybrid systems. Following~\cite{afp:vericomp}, we use modal Kleene algebra, which subsumes the propositional part of dynamic logic, to automatically derive verification conditions for the program flow. Alternatively we also use categorical predicate transformers as formalised in~\cite{afp:transem}. These conditions are entirely about the dynamics that describe the continuous evolution of the hybrid system. The dynamics are formalised with flows and vector fields for systems of ordinary differential equations (ODEs) as in~\cite{ImmlerH12a}. The components support reasoning with vector fields by annotating differential invariants or by providing the solution of the system of ODEs; otherwise, the flow is enough for verification of the continuous evolution. We formalise several rules for derivatives that, when supplied to Isabelle's \isa{auto} method, enhance the automation of the process of discharging proof obligations. In all versions of our verification components we also derive domain specific rules of differential dynamic logic and prove a correctness specification of three hybrid systems using each of our procedures for reasoning with continuous evolutions. In addition to these implementations, for ease of use, we also present a stand alone light-weight variant of the verification components with predicate transformers that does not depend on other AFP entries.

Background information on differential dynamic logic and some of its variants can be found in~\cite{Platzer10,BohrerRVVP17}, the general shallow embedding approach for building verification components with Isabelle can be found in~\cite{ArmstrongGS16}. For more details on modal Kleene algebra see~\cite{DesharnaisS11}; a paper with a detailed overview of the verification components in this entry and the mathematical concepts employed to build them will be available soon on ArXiv.

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
