\documentclass[11pt,a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{mathtools,amsthm,amssymb}
\usepackage{isabelle,isabellesym}

\usepackage[T1]{fontenc}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
\renewcommand{\isastyle}{\isastyleminor}

% LTL Operators

\newcommand{\true}{{\ensuremath{\mathbf{t\hspace{-0.5pt}t}}}}
\newcommand{\false}{{\ensuremath{\mathbf{ff}}}}
\newcommand{\F}{{\ensuremath{\mathbf{F}}}}
\newcommand{\G}{{\ensuremath{\mathbf{G}}}}
\newcommand{\X}{{\ensuremath{\mathbf{X}}}}
\newcommand{\U}{{\ensuremath{\mathbf{U}}}}
\newcommand{\W}{{\ensuremath{\mathbf{W}}}}
\newcommand{\M}{{\ensuremath{\mathbf{M}}}}
\newcommand{\R}{{\ensuremath{\mathbf{R}}}}

% LTL Subformulas

\newcommand{\subf}{\textit{sf}\,}

\newcommand{\sfmu}{{\ensuremath{\mathbb{\mu}}}}
\newcommand{\sfnu}{{\ensuremath{\mathbb{\nu}}}}
\newcommand{\setmu}{\ensuremath{M}}
\newcommand{\setnu}{\ensuremath{N}}
\newcommand{\setF}{\ensuremath{\mathcal{F}}}
\newcommand{\setG}{\ensuremath{\mathcal{G}}}
\newcommand{\setFG}{\ensuremath{\mathcal{F\hspace{-0.1em}G}}}
\newcommand{\setGF}{\ensuremath{\mathcal{G\hspace{-0.1em}F}\!}}

% LTL Functions

\newcommand{\evalnu}[2]{{#1[#2]^\Pi_1}}
\newcommand{\evalmu}[2]{{#1[#2]^\Sigma_1}}
\newcommand{\flatten}[2]{{#1[#2]^\Sigma_2}}
\newcommand{\flattentwo}[2]{{#1[#2]^\Pi_2}}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\title{An Efficient Normalisation Procedure for Linear Temporal Logic: Isabelle/HOL Formalisation}
\author{Salomon Sickert}

\maketitle

\begin{abstract}
In the mid 80s, Lichtenstein, Pnueli, and Zuck proved a classical theorem stating that every formula of Past LTL (the extension of LTL with past operators) is equivalent to a formula of the form $\bigwedge_{i=1}^n \G\F \varphi_i \vee \F\G \psi_i $, where $\varphi_i$ and $\psi_i$ contain only past operators \cite{DBLP:conf/lop/LichtensteinPZ85,XXXX:phd/Zuck86}. Some years later, Chang, Manna, and Pnueli built on this result to derive a similar normal form for LTL \cite{DBLP:conf/icalp/ChangMP92}. Both normalisation procedures have a non-elementary worst-case blow-up, and follow an involved path from formulas to counter-free automata to star-free regular expressions and back to formulas. We improve on both points. We present an executable formalisation of a direct and purely syntactic normalisation procedure for LTL yielding a normal form, comparable to the one by Chang, Manna, and Pnueli, that has only a single exponential blow-up.
\end{abstract}

\tableofcontents

\section{Overview}

This document contains the formalisation of the central results appearing in \cite[Sections 4-6]{XXXX:conf/lics/SickertE20}. We refer the interested reader to \cite{XXXX:conf/lics/SickertE20} or to the extended version \cite{DBLP:journals/corr/abs-2005-00472} for an introduction to the topic, related work, intuitive explanations of the proofs, and an application of the normalisation procedure, namely, a translation from LTL to deterministic automata.

The central result of this document is the following theorem: 

\begin{theorem}
Let $\varphi$ be an LTL formula and let $\Delta_2$, $\Sigma_1$, $\Sigma_2$, and $\Pi_1$ be the classes of LTL formulas from Definition \ref{def:future_hierarchy}. Then $\varphi$ is equivalent to the following formula from the class $\Delta_2$:
\[
\bigvee_{\substack{\setmu \subseteq \sfmu(\varphi)\\\setnu \subseteq \sfnu(\varphi)}} \left( \flatten{\varphi}{\setmu} \wedge \bigwedge_{\psi \in \setmu} \G\F(\evalmu{\psi}{\setnu}) \wedge \bigwedge_{\psi \in \setnu} \F\G(\evalnu{\psi}{\setmu}) \right)
\]
\noindent where $\flatten{\psi}{\setmu}$, $\evalmu{\psi}{\setnu}$, and $\evalnu{\psi}{\setmu}$ are functions mapping $\psi$ to a formula from $\Sigma_2$, $\Sigma_1$, and $\Pi_1$, respectively.
\end{theorem}

\begin{definition}[Adapted from \cite{DBLP:conf/mfcs/CernaP03}]
\label{def:future_hierarchy}
We define the following classes of LTL formulas:
\begin{itemize}
	\item The class $\Sigma_0 = \Pi_0 = \Delta_0$ is the least set containing all atomic propositions and their negations, and is closed under the application of conjunction and disjunction.
	\item The class $\Sigma_{i+1}$ is the least set containing $\Pi_i$ and is closed under the application of conjunction, disjunction, and the $\X$, $\U$, and $\M$ operators.
	\item The class $\Pi_{i+1}$ is the least set containing $\Sigma_i$ and is closed under the application of conjunction, disjunction, and the $\X$, $\R$, and $\W$ operators.
	\item The class $\Delta_{i+1}$ is the least set containing $\Sigma_{i+1}$ and $\Pi_{i+1}$ and is closed under the application of conjunction and disjunction.
\end{itemize}
\end{definition}

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

\bibliographystyle{plainurl}
\bibliography{root}

\end{document}
