\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{amssymb,amsmath}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{Upper Bounding Diameters of State Spaces of Factored Transition Systems}
\author{Friedrich Kurz and Mohammad Abdulaziz}
\maketitle

\begin{abstract}
A {\em completeness threshold} is required to guarantee the completeness of planning as satisfiability, and bounded model checking of safety properties.
One valid completeness threshold is the {\em diameter} of the underlying transition system.
The diameter is the maximum element in the set of lengths of all shortest paths between pairs of states.
The diameter is not calculated exactly in our setting, where the transition system is succinctly described using a (propositionally) factored representation.
Rather, an upper bound on the diameter is calculated compositionally, by bounding the diameters of small abstract subsystems, and then composing those.

We port a HOL4 formalisation of a compositional algorithm for computing a relatively tight upper bound on the system diameter.
This compositional algorithm exploits acyclicity in the state space to achieve compositionality, and it was introduced by Abdulaziz et. al~\cite{icaps2017} (in particular Algorithm~1).
The formalisation that we port is described as a part of another paper by Abdulaziz et. al~\cite{abdulaziz2018formally}, in particular in section~6.
As a part of this porting we developed a libray about transition systems, which shall be of use in future related mechanisation efforts.
\end{abstract}

\tableofcontents

% include generated text of all theories
\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}
