\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{amssymb}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{A Formalization of Declassification with WHAT\&WHERE-Security}
\author{Sylvia Grewe, Alexander Lux, Heiko Mantel, Jens Sauer}
\maketitle

\begin{abstract}
  Research in information-flow security aims at developing methods to
  identify undesired information leaks within programs from private
  sources to public sinks. Noninterference captures this intuition by
  requiring that no information whatsoever flows from private sources
  to public sinks. However, in practice this definition is often too
  strict: Depending on the intuitive desired security policy, the
  controlled declassification of certain private information (WHAT) at
  certain points in the program (WHERE) might not result in an
  undesired information leak.

  We present an Isabelle/HOL formalization of such a security property
  for controlled declassification, namely WHAT\&WHERE-security from
  \cite{scheduler-independent}. The formalization includes
  compositionality proofs for and a soundness proof for a security
  type system that checks for programs in a simple while language with
  dynamic thread creation.

  Our formalization of the security type system is abstract in the
  language for expressions and in the semantic side conditions for
  expressions. It can easily be instantiated with different syntactic
  approximations for these side conditions. The soundness proof of
  such an instantiation boils down to showing that these syntactic
  approximations imply the semantic side conditions.

  This Isabelle/HOL formalization uses theories from the entry
  Strong-Security (see proof document for details).
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
%\input{session}

\section{Preliminary definitions}

\subsection{Type synonyms}

The formalization is parametric in different aspects. Notably, it is
parametric in the security lattice it supports. 

For better readability, we use the following type synonyms in our
formalization (from the entry Strong-Security):

\input{Types.tex}

\section{WHAT\&WHERE-security}

\subsection{Definition of WHAT\&WHERE-security}

The definition of WHAT\&WHERE-security is parametric in a security
lattice (\textit{'d}) and in a programming language (\textit{'com}).

\input{WHATWHERE_Security.tex}


\subsection{Proof technique for compositionality results}

For proving compositionality results for WHAT\&WHERE-security, we formalize
the following ``up-to technique'' and prove it sound:

\input{Up_To_Technique.tex}


\subsection{Proof of parallel compositionality}

We prove that WHAT\&WHERE-security is preserved under composition of
WHAT\&WHERE-secure threads.

\input{Parallel_Composition.tex}


\section{Example language and compositionality proofs}

\subsection{Example language with dynamic thread creation}

As in \cite{scheduler-independent}, we instantiate the language
with a simple while language that supports dynamic thread creation via
a spawn command (Multi-threaded While Language with spawn, MWLs). Note
that the language is still parametric in the language used for Boolean
and arithmetic expressions (\textit{'exp}).

\input{MWLs.tex}

\subsection{Proofs of atomic compositionality results}

We prove for each atomic command of our example programming language
(i.e. a command that is not composed out of other commands) that it is
strongly secure if the expressions involved are indistinguishable for
an observer on security level $d$.

\input{WHATWHERE_Secure_Skip_Assign.tex}

\subsection{Proofs of non-atomic compositionality results}

We prove compositionality results for each non-atomic command of our
example programming language (i.e. a command that is composed out
of other commands): If the components are strongly secure and the
expressions involved indistinguishable for an observer on security
level $d$, then the composed command is also strongly secure.

\input{Language_Composition.tex}

\section{Security type system}

\subsection{Abstract security type system with soundness proof}

We formalize an abstract version of the type system in
\cite{scheduler-independent} using locales
\cite{conf/types/Ballarin03}. Our formalization of the type system is
abstract in the sense that the rules specify abstract semantic side
conditions on the expressions within a command that satisfy for
proving the soundness of the rules. That is, it can be instantiated
with different syntactic approximations for these semantic side
conditions in order to achieve a type system for a concrete language
for Boolean and arithmetic expressions. Obtaining a soundness proof
for such a concrete type system then boils down to proving that the
concrete type system interprets the abstract type system.

We prove the soundness of the abstract type system by simply applying
the compositionality results proven before.

\input{Type_System.tex}

\subsection{Example language for Boolean and arithmetic expressions}

As and example, we provide a simple example language for instantiating the parameter
\textit{'exp} for the language for Boolean and arithmetic expressions
(from the entry Strong-Security).

\input{Expr.tex}

\subsection{Example interpretation of abstract security type system}

Using the example instantiation of the language for Boolean and
arithmetic expressions, we give an example instantiation of our
abstract security type system, instantiating the parameter for domains
\textit{'d} with a two-level security lattice (from the entry Strong-Security).

\input{Domain_example.tex}

\input{Type_System_example.tex}


% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
