\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{url}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}


\begin{document}

\title{Vickrey-Clarke-Groves (VCG) Auctions}
\author{M. B. Caminati\footnote{School of Computer Science, University of Birmingham, UK}\addtocounter{footnote}{-1}
  \and M. Kerber\footnotemark
  \and C. Lange\footnote{Fraunhofer IAIS and University of Bonn, Germany, and School of Computer Science, University of Birmingham, UK}
  \and C. Rowat\footnote{Department of Economics, University of Birmingham, UK}}

\maketitle

\begin{abstract}
A VCG auction (named after their inventors Vickrey, Clarke, and
Groves) is a generalization of the single-good, second price Vickrey
auction to the case of a combinatorial auction (multiple goods, from
which any participant can bid on each possible combination). We
formalize in this entry VCG auctions, including tie-breaking and prove
that the functions for the allocation and the price determination are
well-defined. Furthermore we show that the allocation function
allocates goods only to participants, only goods in the auction are
allocated, and no good is allocated twice. We also show that the price
function is non-negative. These properties also hold for the
automatically extracted Scala code.
\end{abstract}


\tableofcontents

\section{Introduction}
An auction mechanism is mathematically represented through a pair of
functions $(a, p)$: the first describes how some given goods at stake
are allocated among the bidders (also called participants or agents),
while the second specifies how much each bidder pays following this
allocation.  Each possible output of this pair of functions is
referred to as an outcome of the auction. Both functions take the same
argument, which is another function, commonly called a bid vector $b$;
it describes how much each bidder values the possible outcomes of the
auction. This valuation is usually expressed through money.  In this
setting, some common questions are the study of the quantitative and
qualitative properties of a given auction mechanism (e.g., whether it
maximizes some relevant quantity, such as revenue, or whether it is
efficient, that is, whether it allocates the item to the bidder who
values it most), and the study of the algorithms running it (in
particular, their correctness).

A VCG auction (named after their inventors Vickrey, Clarke, and
Groves) is a generalization of the single-good, second price Vickrey
auction to the case of a combinatorial auction (multiple goods, from
which any participant can bid on each possible combination). We
formalize in this entry VCG auctions, including tie-breaking and prove
that the functions $a$ and $p$ are well-defined. Furthermore we show
that the allocation function $a$ allocates goods only to participants,
only goods in the auction are allocated, and no good is allocated
twice. Furthermore we show that the price function $p$ is
non-negative. These properties also hold for the automatically
extracted Scala code. For further details on the formalization, see
\cite{ec15}. For background information on VCG auctions, see \cite{cramton}.

  
The following files are part of the Auction Theory Toolbox
(ATT)~\cite{github} developed in the ForMaRE project~\cite{formare}.
The theories \texttt{CombinatorialAuction.thy},
\texttt{StrictCombinatorialAuction.thy} and
\texttt{UniformTieBreaking.thy} contain the relevant definitions and
theorems; \texttt{CombinatorialAuctionExamples.thy} and
\texttt{CombinatorialAuctionCodeExtraction.thy} present simple helper
definitions to run them on given examples and to export
them to the Scala language, respectively; \texttt{FirstPrice.thy}
shows how easy it is to adapt the definitions to the first price
combinatorial auction.  The remaining theories contain more general
mathematical definitions and theorems.


\subsection{Rationale for developing set theory as replacing one bidder in a second price auction}

Throughout the whole ATT, there is a duality in the way mathematical
notions are modeled: either through objects typical of lambda calculus
and HOL (lambda-abstracted functions and lists, for example) or
through objects typical of set theory (for example, relations,
intersection, union, set difference, Cartesian product).

This is possible because inside HOL, it is possible to model a
simply-typed set theory which, although quite restrained if compared
to, e.g., ZFC, is powerful enough for many standard mathematical
purposes.

ATT freely adopts one approach, the other, or a mixture thereof, depending on technical and
expressive convenience.
A technical discussion of this topic can be found in~\cite{cicm2014}.

\subsection{Bridging}

One of the differences between the approaches of functional
definitions on the one hand and classical (often set-theoretical)
definitions on the other hand is that, commonly (although not always),
the first approach is better suited to produce Isabelle/HOL
definitions which are computable (typically, inductive definitions);
while the definitions from the second approach are often more general
(e.g., encompassing infinite sets), closer to pen-and-paper
mathematics, but also not computable.  This means that many theorems
are proved with respect to definitions of the second type, while in
the end we want them to apply to definitions of the first type,
because we want our theorems to hold for the code we will be actually
running.  Hence, bridging theorems are needed, showing that, for the
limited portions of objects for which we state both kinds of
definitions, they are the same.

\subsection{Main theorems}

The main theorems about VCG auctions are:
\begin{description}
  \item[the definiteness theorem:] our definitions grant that there is exactly one solution; this is 
    ensured by \texttt{vcgaDefiniteness}.
  \item[PairwiseDisjointAllocations:] no good is allocated to more than one participant.
  \item[onlyGoodsAreAllocated:] only the actually available goods are allocated.
  \item[the adequacy theorem:] the solution provided by our algorithm is indeed the one prescribed by
    standard pen-and-paper definition.
  \item[NonnegPrices:] no participant ends up paying a negative price 
    (e.g., no participant receives money at the end of the auction).
  \item[Bridging theorems:] as discussed above, such theorems permit to apply the theorems in this 
    list to the executable code Isabelle generates.
\end{description}

\subsection{Scala code extraction}

Isabelle permits to generate, from our definition of VCG, Scala code
to run any VCG auction.  Use
\texttt{CombinatorialAuctionCodeExtraction.thy} for this. This code is
in the form of Scala functions which can be evaluated on any input
(e.g., a bidvector) to return the resulting allocation and prices.

To deploy such functions use the provided Scala
wrapper (taking care of the output and including sample inputs).  In
order to do so, you can evaluate inside Isabelle/JEdit the file
\texttt{CombinatorialAuctionCodeExtraction.thy} (position the cursor on its
last line and wait for Isabelle/JEdit to end all its processing).
This will result in the file
\texttt{/dev/shm/VCG-withoutWrapper.scala}, which can be automatically
appended to the wrapper by running the shell script at
the end of \texttt{CombinatorialAuctionCodeExtraction.thy}. For details of how to run the Scala code see
\url{http://www.cs.bham.ac.uk/research/projects/formare/vcg.php}.


% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
