\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym,fullpage}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}


\begin{document}

\title{Partial Semigroups and Convolution Algebras}
\author{Brijesh Dongol, Victor B F Gomes, Ian J Hayes and Georg
  Struth}


\maketitle

\begin{abstract}
  Partial Semigroups are relevant to the foundations of quantum
  mechanics and combinatorics as well as to interval and separation
  logics. Convolution algebras can be understood either as algebras of
  generalised binary modalities over ternary Kripke frames, in
  particular over partial semigroups, or as algebras of
  quantale-valued functions which are equipped with a
  convolution-style operation of multiplication that is parametrised
  by a ternary relation.  Convolution algebras provide algebraic
  semantics for various substructural logics, including categorial,
  relevance and linear logics, for separation logic and for interval
  logics; they cover quantitative and qualitative applications. These
  mathematical components for partial semigroups and convolution
  algebras provide uniform foundations from which models of
  computation based on relations, program traces or pomsets, and
  verification components for separation or interval temporal logics
  can be built with little effort.
\end{abstract}

\tableofcontents

\section{Introductory Remarks}

These mathematical components supply formal proofs for two articles on
\emph{Convolution Algebras}~\cite{DongolHS17} and \emph{Convolution as
  a Unifying Concept}~\cite{DongolHS16}.  They are sparsely documented
and referenced; additional information can be found in these articles,
and in particular the first one.

The approach generalises previous Isabelle components for covolution
algebras that were intended for separation logic and used partial
abelian semigroups and monoids for modelling store-heap
pairs~\cite{DongolGS15}.  Due to the applications in separation logic,
a detailed account of cancellative and positive partial abelian
monoids has been included, as these structures characterise the heap
succinctly.  Isabelle verification components based on this
approach will be submitted as a separate AFP entry.
 
Our article on convolution algebras~\cite{DongolHS17} provides a
detailed account of convolution-based semantics for
Halpern-Shoham-style interval logics~\cite{HalpernS91,Venema91},
interval temporal logics~\cite{Moszkowski00} and duration
calculi~\cite{ZhouH04} based on partial monoids.  While general
approaches, including modal algebras over semi-infinite intervals, are
supported by the mathematical components provided, additional work on
store models and assignments of variables to values is needed in order
to build verification components for such interval logics.

Convolution-based liftings of partial semigroups of graphs and partial
orders allow formalisations of models of true concurrency such as
pomset languages and concurrent Kleene algebras~\cite{HoareMSW11} in
Isabelle, too.  An AFP entry for these is in preparation.

In all these approaches, the main task is to construct suitable
partial semigroups or monoids of the computational models intended,
for instance, closed intervals over the reals under fusion product,
unions of heaplets (i.e. partial functions) provided their domains are
disjoint, disjoint unions of graphs as parallel products.  Our
approach then allows a generic lifting to convolution algebras on
suitable function spaces with algebraic properties, for instance of
heaplets to the assertion algebra of separation logic with separating
conjunction as convolution~\cite{DongolGS15,DongolHS16}, or of
intervals to algebraic counterparts of interval temporal logics or
duration calculi with the chop operation as
convolution~\cite{DongolHS17}.  We believe that this general
construction supports other applications as well---qualitative and
quantitative ones.

We would like to thank Alasdair Armstrong for his help with some
Isabelle proofs and Tony Hoare for many discussions that helped us
shaping the general approach.

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
